---
title: "maxa3v3"
output: html_document
date: "2024-12-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(janitor)
library(formattable)
library(sf)
library(ggplot2)
library(tmap)

```


```{r read and clean calls}
#import csv cleaned for NYU and CU zip codes
calls <- read.csv("311_Noise_Complaints.csv")
# Convert "Created.Date" to Date format and extract the year
calls <- calls %>%
  mutate(Created.Date = as.Date(Created.Date, format = "%m/%d/%Y"),
         Year = as.numeric(format(Created.Date, "%Y")))

# Specify the noise complaint types to keep
descriptors_to_keep <- c(
  "Banging/Pounding", "Loud Music/Party", "Loud Talking", "Loud Television",
  "Noise: Alarms", "Noise: Construction Before/After Hours (NM1)",
  "Noise: Construction Equipment (NC1)", "Noise: Jack Hammering (NC2)",
  "Noise: Loud Music/Daytime (Mark Date and Time) (NN1)",
  "Noise: Loud Music/Nighttime(Mark Date and Time) (NP1)",
  "People Created Noise"
)

#Filter the dataset by the complaint types
calls_filtered <- calls %>%
  filter(Descriptor %in% descriptors_to_keep) %>%
  filter(!is.na(Latitude) & !is.na(Longitude)) %>%  #count NAs
  clean_names()

#convert to sf
calls_sf <- st_as_sf(calls_filtered, coords = c("longitude", "latitude"), remove = FALSE, crs = 2236)
```


```{r Wash Sq Park buffer}
#COMMENTING THIS OUT BC WE ARE NO LONGER USING THIS BUFFER 

# Define your point as an sf object
point <- st_as_sf(data.frame(id = 1,
                             lon = -73.9999069,
                             lat = 40.7308878),
                  coords = c("lon", "lat"), crs = 2236)
                  

point_proj <- st_transform(point, crs = 2236)

buffer <- st_buffer(point_proj, dist = 1200)

st_crs(calls_sf) == st_crs(buffer) #test crs

plot(st_geometry(buffer), col = "white", border = "blue")  # Plot buffer
plot(st_geometry(calls_sf), add = TRUE, col = "red")      # Add calls points

# Filter calls_sf points that are within the buffer
data_within_buffer <- calls_sf %>%
  filter(st_within(geometry, buffer, sparse = FALSE) %>% rowSums() > 0)

# #MAPPING

# allcallsmap <- ggplot() +
#   geom_sf(data = calls_sf, aes(color = "311 Calls"), alpha = 0.5, size = 0.5) +
#   scale_color_manual(values = "red") +
#   theme_minimal() +
#   labs(title = "311 Calls Heatmap (Transparency Effect)",
#        color = NULL)
#
# print(allcallsmap)

heatmap <- ggplot() +
  geom_sf(data = buffer, fill = NA, color = "blue") +  # Buffer boundary
  geom_sf(data = data_within_buffer, aes(color = "311 Calls"), alpha = 0.5, size = 0.5) +
  scale_color_manual(values = "red") +
  theme_void() +
  labs(title = "311 Noise Calls Heatmap (Transparency Effect)",
       subtitle = "Within 1200m buffer",
       color = NULL)

print(heatmap)

write_sf(buffer, "1200.radius.buffer.geojson")
```

```{r justfix props w BBLs}

# ---------------------------------------------------------------------------------------------------
#ignore if you use file from drive - THIS IS HOW TO GET TO SHAPEFILE I READ IN BELOW (we can avoid using entire mappluto file now)

# #mappluto shapefile
# mappluto <- st_read("nyc_mappluto_24v3_1_fgdb/MapPLUTO24v3_1.gdb")
# 
# head(mappluto)
# 
# mappluto_clean <- mappluto %>%
#   select(Address, APPBBL)
# 
# #writing this to git for easier access - but it will be in drive and ignored on github
# st_write(mappluto_clean, "mappluto_bbl_address.geojson")
# ---------------------------------------------------------------------------------------------------

# mappluto <- st_read("mappluto_bbl_address.geojson") #from google drive /PLUTO
# 
# #nyu landlord properties (justfix)
# justfix_nyu_props <- read.csv("nyuprops_who_owns_what_export.csv") #from github or drive /justfix
# 
# # Filter the BBL polygons based on your property BBLs
# properties_sf_bbl <- mappluto[mappluto$APPBBL %in% justfix_nyu_props$bbl, ]
# 
# #write out shapefiles of these properties
# # st_write(properties_sf_bbl, "nyuprops_bbl_address.geojson")
# 
# 
# # View the result
# plot(st_geometry(properties_sf_bbl), col = "purple")

```

```{r res hall props}
#mappluto shapefile

mappluto_clean <- read_sf("mappluto_bbl_address.geojson") %>%
  clean_names()

# Original address table
nyu_residence_halls <- data.frame(
  Building = c(
    "Brittany Hall", "Founders Hall", "Lipton Hall", "Res College at Paulson",
    "Rubin Hall", "Weinstein Hall", "Alumni Hall", "Broome Street",
    "Carlyle Court", "Coral Tower", "Gramercy Green", "Greenwich Hall",
    "Lafayette Hall", "Palladium Hall", "Second Street", "Seventh Street",
    "Sixth Street", "Third North", "University Hall"
  ),
  Address = c(
    "55 East 10th Street", "120 East 12th Street", "33 Washington Square",
    "181 Mercer Street", "35 Fifth Street", "5 University Place",
    "33 Third Avenue", "400 Broome Street", "25 Union Square West",
    "129 Third Avenue", "310 3rd Avenue", "636 Greenwich Street",
    "80 Lafayette Street", "140 East 14th Street", "1 East 2nd Street",
    "40 East Seventh Street", "200 East 6th St", "75 Third Avenue",
    "110 East 14th Street"
  )
)

# Add BBL and Add 2 columns
nyu_residence_halls <- nyu_residence_halls %>%
  mutate(
    BBL = c(
      56230, 55648, 55224, 52466, 4613, 5484, 4651, 4811, 84322, 8961,
      87944, 60346, 17220, 55922, 4589, 46218, 4616, 4671, 55912
    ),
    Add_2 = c(
      "787 BROADWAY", "120 EAST 12 STREET", "35 WASHINGTON SQUARE W", 
      "100 BLEECKER STREET", "25 COOPER SQUARE", "5 UNIVERSITY PLACE", 
      "31 3 AVENUE", "400 BROOME STREET", "21 UNION SQUARE", "125 3 AVENUE", 
      "318 3 AVENUE", "636 GREENWICH STREET", "76 LAFAYETTE STREET", "126 EAST 14 STREET", 
      "1 EAST 2 STREET", "38 EAST 7 STREET", "35 COOPER SQUARE", 
      "67 3 AVENUE", "106 EAST 14 STREET"
    )
  )

# Add the Manhattan borough code ("1") as a prefix to BBL
nyu_residence_halls <- nyu_residence_halls %>%
  mutate(
    BBL = paste0("1", sprintf("%09d", as.numeric(BBL))) # Ensure numeric BBL and pad to 9 digits
  )

# Normalize addresses in both datasets
nyu_residence_halls <- nyu_residence_halls %>%
  mutate(Add_2 = str_to_upper(Add_2)) %>%  # Convert to uppercase
  clean_names()

str(nyu_residence_halls$bbl)
str(mappluto_clean$bbl)

# mappluto_clean <- mappluto_clean %>%
#   mutate(address = str_trim(str_to_upper(address)))   


# View the updated dataset
print(nyu_residence_halls)

# # Geocode the addresses to get APPBBL (replace with real APPBBL values if known)
# geocoded_halls <- data.frame(
#   address = nyu_residence_halls$address,
#   APPBBL = c(
#     # Add APPBBL values manually or use a geocoding step here
#   )
# )

### troubleshooting addresses

# Look at some mappluto addresses
head(mappluto_clean)


# Filter the mappluto dataset
nyu_residence_halls_geom <- mappluto_clean %>%
  filter(address %in% nyu_residence_halls$add_2) %>%
  select(-appbbl)


# Check the result
summary(nyu_residence_halls_geom)


# View the result
plot(st_geometry(nyu_residence_halls_geom), col = "purple")

# Now write the shapefile
st_write(nyu_residence_halls_geom, "residence_halls.shp", append = TRUE)

```


```{r bbl props buffer}
# read margots QGIS shapefile
qgis_halls <- read_sf("~/Desktop/GitHub/nyu-noise-v5/NYU_MN_Residence_Halls_CleanedGeo.geojson")

# Create a 100-foot buffer around the properties
buffer_100ft <- st_buffer(qgis_halls, dist = 100) %>%
  st_transform(crs = st_crs(qgis_halls))

# check crs
st_crs(buffer_100ft)
st_crs(nyu_residence_halls_geom)

# Plot the result: Properties (purple points) and 100ft buffer (blue)
plot(st_geometry(buffer_100ft), border = "blue", main = "NYU Properties and 100ft Buffer")
plot(st_geometry(nyu_residence_halls_geom), add = TRUE, col = "purple", pch = 16)
plot(st_geometry(data_within_buffer), add = TRUE, col = "pink", pch = 16)

# write out shapefiles of buffers
st_write(buffer_100ft, "BUFFERS_NYU_MN_Residence_Halls_CleanedGeo.geojson", append = TRUE)

# st_crs(buffer)
# nyu_properties <- nyu_properties %>% st_transform(4326)
```

```{r properties with buffer}

# # Ensure CRS compatibility
st_crs(data_within_buffer)
summary(data_within_buffer$longitude)  # Should be around -74 for NYC
summary(data_within_buffer$latitude)   # Should be around 40.7 for NYC

st_crs(nyu_properties)
summary(nyu_properties$longitude)  # Should be around -74 for NYC
summary(nyu_properties$latitude)   # Should be around 40.7 for NYC

data_within_buffer <- st_transform(data_within_buffer, crs = st_crs(nyu_properties))

# # having crs issues :/ should probably figure this out in QGIS
# calls_and_props <- st_join(calls_sf, properties_sf_bbl)
# plot(st_geometry(calls_and_props))

# Create the heatmap
heatmap <- ggplot() +
  # Plot 311 Calls with transparency (alpha)
  geom_sf(data = data_within_buffer, aes(color = "Noise Complaints"), alpha = 0.5, size = 0.5) +
  
  # Plot Buffer layer with light blue semi-transparent fill and no outline
  geom_sf(data = buffer_100ft, fill = "100ft Buffer", color = NA, border = "blue", alpha = 0.5) +
  
  # Plot properties with purple outline and no fill (transparent)
  geom_sf(data = nyu_residence_halls_geom, aes(color = "NYU Properties"), fill = NA, alpha = 1, size = 2) +
  
  # Customize theme and labels
  theme_minimal() +
  labs(title = "Noise Calls Heatmap",
       subtitle = "Within 1200ft of WSP, 100ft buffers around properties",
       color = NULL) +
  
  #Add a custom color palette for clarity
  scale_color_manual(values = c("Noise Complaints" = "red", "NYU Properties" = "purple", "100ft Buffer" = "lightblue"))


print(heatmap)

st_write(data_within_buffer, "calls_within_buffer_2263.geojson")

```

## MARGOT CODE #####
## BUILDINGS WITH CALLS SF TO CSV##

# Step 1: Read the GeoJSON file
calls_with_buildings <- st_read("Calls_Associated_with_Buildings.geojson")

# Step 2: Remove geometry if not needed (optional)
# Convert to a data frame, drop the geometry column, and rename 'address_2' to 'nyu_address'
calls_with_buildings_df <- calls_with_buildings %>%
  st_drop_geometry() %>%
  rename(nyu_address = address_2)  # Rename address_2 to nyu_address

# Step 3: Save the data frame as a CSV file
write.csv(calls_with_buildings_df, "Calls_Associated_with_Buildings.csv", row.names = FALSE)

# Alternatively, using readr for better control over formatting
# readr::write_csv(calls_with_buildings_df, "Calls_Associated_with_Buildings.csv")

# Read the dataset
calls_associated_buildings <- read.csv("Calls_Associated_with_Buildings.csv")

# Select only the desired columns and save to a new object
calls_associated_buildings_clean <- calls_associated_buildings %>%
  select(unique_key, complaint_, descriptor, incident_a, street_nam, location, year, address_2, distance)

# Rename the column 'address' to 'nyu_address'
calls_associated_buildings_clean <- calls_associated_buildings_clean %>%
  rename(nyu_address = address_2)

# View the resulting dataset
head(calls_associated_buildings_clean)

write.csv(calls_associated_buildings_clean, "Calls_Associated_with_Buildings_Clean.csv", row.names = FALSE)


# SUMMARIZE ##
# Group by nyu_address and descriptor, then count the total number
summary_by_address_descriptor <- calls_associated_buildings_clean %>%
  group_by(nyu_address, descriptor) %>%
  summarise(total_calls = n(), .groups = "drop") %>%
  pivot_wider(names_from = descriptor, values_from = total_calls, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) %>% # Add row-wise totals
  bind_rows(
    summarise_all(., ~ if(is.numeric(.)) sum(.) else "Total") # Add total row
  )

# View the reshaped summary
print(summary_by_address_descriptor)

# Save the summary to a CSV file
write.csv(summary_by_address_descriptor, "Summary_by_Address_Descriptor.csv", row.names = FALSE)

##SUMMARY 2##

# Group by nyu_address and year, then count the total number of calls
summary_by_address_year <- calls_associated_buildings_clean %>%
  group_by(nyu_address, year) %>%
  summarise(total_calls = n(), .groups = "drop") %>%
  pivot_wider(names_from = nyu_address, values_from = total_calls, values_fill = 0) %>%
  mutate(Total = rowSums(across(where(is.numeric)))) # Add row-wise totals

# Add a bottom row for total calls across all years and addresses
summary_by_address_year <- summary_by_address_year %>%
  bind_rows(
    summarise_all(., ~ if(is.numeric(.)) sum(.) else "Total")
  )
# Save the summary to a CSV file
write.csv(summary_by_address_year, "Summary_by_Address_Year.csv", row.names = FALSE)

## SUMMARY 2 MODIFIED WITHOUT TOTAL COLUMN AND ROW ##

# Group by nyu_address and year, then count the total number of calls
summary_by_address_year2 <- calls_associated_buildings_clean %>%
  group_by(nyu_address, year) %>%
  summarise(total_calls = n(), .groups = "drop") %>%
  pivot_wider(names_from = nyu_address, values_from = total_calls, values_fill = 0)

# Save the summary to a CSV file
write.csv(summary_by_address_year2, "Summary_by_Address_Year2.csv", row.names = FALSE)






